### 2021-03-12 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Finished proofs for SH work. Removed Tangent space stuff from CH4, made appendix. Need to address comments.

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and are given by \bstodo{Add citation for associated legendre polys}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; \chat \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x) \\
	P^{-m}_l(x) &:= (-1)^m \; \ctilde \; \Plm(x),
}
for $m = 0, 1, 2, \dots, l$ where
\bseqn{
	\chat &:= {(l+m)! \over 2^m \; l!} \\
	\ctilde &:= {(l-m)! \over (l+m)!}
}
and $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0! = 1$. These relationships for the Jacobi polynomials and associated Legendre polynomials allow us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.

On that note, we can now write down the spherical harmonics. Let $(x,y,z) \in \Omega$. It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$). Throughout this document we will use the convention that the spherical coordinate angles be defined by
\bseqn{
	x &= \sinphi \costheta = \rho(z) \costheta \\
	y &= \sinphi \sintheta = \rho(z) \sintheta \\
	z &= \cosphi.
}
where $\rho(z) := \sqrt{1-z^2}$. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
\bseqnnumber{
	\Ylmfull &:= \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \: \eimtheta \Plm (\cosphi) \nonumber \\
	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
}
for $0 \le |m| \le l, \, l \in \No$, where
\bseqnnumber{
	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
		\begin{cases} 
			(-1)^m \: \chat &\text{if } m \ge 0 \\
			\chatmod &\text{if } m < 0
		\end{cases}. \label{eqn:clmdef}
} 
The spherical harmonics as defined are then orthonormal with respect to the inner product
\bseqn{
	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
	& \quad = \delta_{l, l'} \: \delta_{m, m'}
}
where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$. Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m \in \Z$. Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

@@

Using \bsrefeqns{eqn:shrecx}{eqn:shrecz}, we have that the Jacobi matrices take the following block-tridiagonal form (i.e. have block-bandwidths $(1,1)$):
\bseqn{
	J_x &= 
		\begin{pmatrix}
			B_{x,0} & A_{x,0} & & & & \\
			C_{x,1} & B_{x,1} & A_{x,1} & & & \\
			& C_{x,2} & B_{x,2} & A_{x,2}  & & & \\
			& & C_{x,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_y &= 
		\begin{pmatrix}
			B_{y,0} & A_{y,0} & & & & \\
			C_{y,1} & B_{y,1} & A_{y,1} & & & \\
			& C_{y,2} & B_{y,2} & A_{y,2}  & & & \\
			& & C_{y,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_z &= 
		\begin{pmatrix}
			B_{z,0} & A_{z,0} & & & & \\
			C_{z,1} & B_{z,1} & A_{z,1} & & & \\
			& C_{z,2} & B_{z,2} & A_{z,2}  & & & \\
			& & C_{z,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
$J_x$ has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,3} & 0 & \alpha_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \alpha_{l,l,3} & 0 & \alpha_{l,l,4} \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{x,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,2} & & \\
			0 & \ddots & \\
			\alpha_{l,-l+2,1} & \ddots & \alpha_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \alpha_{l,l,1}
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
$J_y$ also has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,3} & 0 & \beta_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \beta_{l,l,3} & 0 & \beta_{l,l,4} \\
		\end{pmatrix} \in \C^{(2l+1)\times(2l+3)}, \\
	B_{y,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,2} & & \\
			0 & \ddots & \\
			\beta_{l,-l+2,1} & \ddots & \beta_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \beta_{l,l,1}
		\end{pmatrix} \in \C^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
Finally, $J_z$ has subblock-bandwidths $(1,1)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{z,l} &:= 
		\begin{pmatrix}
			0 & \gamma_{l,-l,2} & 0 & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 &  \gamma_{l,-l,2} & 0 \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{z,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{z,l} &:= 
		\begin{pmatrix}
			0 & & \\
			\gamma_{l,-l+1,1} & \ddots & \\
			0 & \ddots & 0 \\
			& \ddots & \gamma_{l,l-1,1} \\
			& & 0
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}

@@

Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we set
\bseqnnumber{
	\Dlt = 
		\begin{pmatrix}
			\hat{A}^{x,y}_l & \zeromat{(2l+3)\times(2l+1)}
		\end{pmatrix} \in \R^{(2l+3)\times3(2l+1)} \label{eqn:Dltdef}
}
where $\zeromat{(2l+3)\times(2l+1)}$ the zero matrix in $\R^{(2l+3)\times(2l+1)}$, and $\hat{A}^{x,y}_l \in \R^{(2l+3)\times2(2l+1)}$ is the left inverse of the matrix $\begin{pmatrix} A^x_l \\ A^y_l \end{pmatrix}$, given by
\bseqn{
	\hat{A}^{x,y}_l = 
		\begin{pmatrix}
			\frac{1}{2\alpha_{l,-l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} & 0 & \hdots & 0 \\
			& \ddots & & & & \ddots & & \\
			& & \ddots & & & & \ddots & \\
			& & & \frac{1}{2\alpha_{l,l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} \\
			0 & \hdots & \frac{1}{2\alpha_{l,l-1,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l-1,4}} & 0 \\
			0 & \hdots & 0 & \frac{1}{2\alpha_{l,l,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l,4}}
		\end{pmatrix}.
}
For $l = 0$ we set
\bseqnnumber{
	D^T_0 = 
		\begin{pmatrix}
			\frac{1}{2\alpha_{0,0,3}} & \frac{1}{2\beta_{0,0,3}} & 0 \\
			0 & 0 & \frac{1}{\gamma_{0,0,2}} \\
			\frac{1}{2\alpha_{0,0,4}}&  \frac{1}{2\beta_{0,0,4}} & 0 
		\end{pmatrix}. \label{eqn:Dzerotdef}
}

It will be useful for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
\begin{definition}\label{def:clenshawmats}
	The matrices $-\Dlt (B_l-G_l(x,y,z))$, $\Dlt C_l$ for $l \in \No$ defined via \bsrefeqns{eqn:clenshawmats1}{eqn:Dzerotdef} are called the \textbf{recurrence coefficient matrices} for a given family of multidimensional orthogonal polynomials.
\end{definition}

@@

\bstodoinline{obtaining coeffs/ transform}
In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations. Naturally, we of course need a way to obtain the coefficients $f_{l,m}$. The coefficients can be calculated via the integral
\bseqn{
	f_{l,m} = \int_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
}
using the orthonormality of the spherical harmonics. Methods to calculate these coefficients exist, known as spectral transforms, and are well established for the spherical harmonics (see e.g. \cite{mohlenkamp1999fast, suda2002fast}).

@@

Let's explain what we mean. Suppose $v : \Omega \to \C$ is a function, and we encounter the problem of finding the coefficients of the expansion for $v(x,y,z) f(x,y,z)$. We wish to therefore find an operator $V$ for $v$ so that
\bseqn{
	v(x,y,z) f(x,y,z) = v(x,y,z) \fvec^\top \SHvec(x,y,z) = (V \; \fvec)^\top \SHvec(x,y,z),
}
i.e. $V \; \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto v(x,y,z) f(x,y,z)$ in the spherical harmonic basis.

@@

%%% TANGENT SPACE ----- VECTOR SHs
\section{Vector spherical harmonics as orthogonal vectors in three variables}

@@

As for the scalar case, these recurrences above lead us to Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
\bseqnnumber{
	\VSHvecfull = 
		\begin{pmatrix} 
			\VSHvecfull_0 \\ \VSHvecfull_1 \\ \vdots 
		\end{pmatrix}, \quad \text{where } \quad 
	\VSHvecfulll = 
		\begin{pmatrix} 
			(\gradY^{-l}_l)^\top \\ (\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top \\ (\gradpY^{l}_l)^\top
		\end{pmatrix} \in \C^{2(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHfulldef}
}

@@

			\mathcal{A}_{l,-l,5} & 0_2 & \mathcal{A}_{l,-l,6} & & \\

@@

			& & \mathcal{A}_{l,l,5} & 0_2 & \mathcal{A}_{l,l,6} \\

@@

			0_2 & \mathcal{A}_{l,-l,4} & \\
			\mathcal{A}_{l,-l+1,3} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{A}_{l,l-1,4} \\
			& & \mathcal{A}_{l,l,3} & 0_2

@@

			0 & A_{l,m,j}^*
		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad

@@

			A_{l,m,j}^* & 0
		\end{pmatrix} \: \text{for } j = 3,4.

@@

			\mathcal{B}_{l,-l,5} & 0_2 & \mathcal{B}_{l,-l,6} & & \\

@@

			& & \mathcal{B}_{l,l,5} & 0_2 & \mathcal{B}_{l,l,6} \\

@@

			0_2 & \mathcal{B}_{l,-l,4} & \\
			\mathcal{B}_{l,-l+1,3} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{B}_{l,l-1,4} \\
			& & \mathcal{B}_{l,l,3} & 0_2

@@

			0 & B_{l,m,j}^*
		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad

@@

			B_{l,m,j}^* & 0
		\end{pmatrix} \: \text{for } j = 3,4.

@@

			0 & 0 & \Gamma_{l,-l,3} \\
			& & & \Gamma_{l,-l,3}^* \\

@@

			& & & & & \Gamma_{l,l,3} \\
			& & & & & & \Gamma_{l,l,3}^*& 0 & 0 \\

@@

			0 & \Gamma_{l,-l,2} \\
			\Gamma_{l,-l,2}^* & 0 \\

@@

			& & & 0 & \Gamma_{l,l,2} \\
			& & & \Gamma_{l,l,2}^* & 0 \\

@@

			& \Gamma_{l,-l+1,1}^* \\

@@

			& & & & \Gamma_{l,l-1,1}^* \\

@@

Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
\bseqn{
	\uvec(x,y,z) = u_\varphi(x,y,z) \: \phivec + u_\theta(x,y,z) \: \thetavec, \quad (x,y,z) \in \Omega
}
for some scalar functions $u^\varphi, u^\theta$ where $\phivec, \thetavec$ are the standard non-radial unit basis vectors on the sphere defined by $\phivec := (\cos \theta \sin \varphi, \sin \theta \sin \varphi, \cos \varphi)^\top$, $\thetavec := (-\sin \theta, \cos \theta, 0)^\top$. Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}. We can then approximate this function by its expansion in the VSH basis, i.e. for some large enough $N \in \N$:
\bseqn{
	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big].
}

>>>>

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Jacobi polynomial: \bstodo{Add citation for associated legendre polys}
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; {(l+m)! \over 2^m \; l!} \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x),
}
for $m = 0,1,2,\dots,l$. It is also standard to include a further definition for the case of $m$ negative -- that is we also define:
\bseqn{
	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x)
}
for $m = -l, \dots, -1$. Here, $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0! = 1$. Although referred to as the associated Legendre \enquote{polynomials}, they are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions instead. Moreover, it should be noted too that the inclusion of the $(-1)^m$ factor in the definition for $m > 0$, known as the \textit{Condon-Shortley phase} in physics, is sometimes omitted.

These relationships for the Jacobi polynomials and associated Legendre polynomials allow us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.

On that note, we can now write down the spherical harmonics. Let $(x,y,z) \in \Omega$. It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$). Throughout this document we will use the convention that the spherical coordinate angles be defined by
\bseqn{
	x &= \sinphi \costheta = \rho(z) \costheta \\
	y &= \sinphi \sintheta = \rho(z) \sintheta \\
	z &= \cosphi.
}
where $\rho(z) := \sqrt{1-z^2}$. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
\bseqnnumber{
	\Ylmfull &:= \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \: \eimtheta \Plm (\cosphi) \nonumber \\
	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
}
for $0 \le |m| \le l, \, l \in \No$, where
\bseqnnumber{
	\clm := \Bigg({(2l+1) \: (l-m)! \over 4\pi \: (l+m)!}\Bigg)^\half \: {(l+m)! \over 2^{\abs{m}} \; l!} \:
		\begin{cases} 
			(-1)^m &\text{if } m \ge 0 \\
			1 &\text{if } m < 0
		\end{cases}. \label{eqn:clmdef}
} 
The spherical harmonics as defined are then orthonormal with respect to the inner product
\bseqn{
	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
	& \quad = \delta_{l, l'} \: \delta_{m, m'}
}
where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$. Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m \in \Z$. Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

@@

Using \bsrefeqns{eqn:shrecx}{eqn:shrecz}, we have that the Jacobi matrices take the following block-tridiagonal form (i.e. have block-bandwidths $(1,1)$):
\bseqn{
	J_x &= 
		\begin{pmatrix}
			B_{x,0} & A_{x,0} & & & & \\
			C_{x,1} & B_{x,1} & A_{x,1} & & & \\
			& C_{x,2} & B_{x,2} & A_{x,2}  & & & \\
			& & C_{x,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_y &= 
		\begin{pmatrix}
			B_{y,0} & A_{y,0} & & & & \\
			C_{y,1} & B_{y,1} & A_{y,1} & & & \\
			& C_{y,2} & B_{y,2} & A_{y,2}  & & & \\
			& & C_{y,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_z &= 
		\begin{pmatrix}
			B_{z,0} & A_{z,0} & & & & \\
			C_{z,1} & B_{z,1} & A_{z,1} & & & \\
			& C_{z,2} & B_{z,2} & A_{z,2}  & & & \\
			& & C_{z,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
$J_x$ has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,3} & 0 & \alpha_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \alpha_{l,l,3} & 0 & \alpha_{l,l,4} \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{x,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,2} & & \\
			0 & \ddots & \\
			\alpha_{l,-l+2,1} & \ddots & \alpha_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \alpha_{l,l,1}
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
$J_y$ also has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,3} & 0 & \beta_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \beta_{l,l,3} & 0 & \beta_{l,l,4} \\
		\end{pmatrix} \in \C^{(2l+1)\times(2l+3)}, \\
	B_{y,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,2} & & \\
			0 & \ddots & \\
			\beta_{l,-l+2,1} & \ddots & \beta_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \beta_{l,l,1}
		\end{pmatrix} \in \C^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
Finally, $J_z$ has subblock-bandwidths $(1,1)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{z,l} &:= 
		\begin{pmatrix}
			0 & \gamma_{l,-l,2} & 0 & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 &  \gamma_{l,l,2} & 0 \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{z,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{z,l} &:= 
		\begin{pmatrix}
			0 & & \\
			\gamma_{l,-l+1,1} & \ddots & \\
			0 & \ddots & 0 \\
			& \ddots & \gamma_{l,l-1,1} \\
			& & 0
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}

@@

Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we can set
\bseqnnumber{
	\Dlt = 
		\begin{pmatrix}
			& & & & & \vec{\eta}^\top_{0} \\
			0 & & & 0 & & & {1 \over \gamma_{l,-l,2}} & \\
			& \ddots & & & \ddots & & & \ddots \\
			& & 0 & & & 0 & & & {1 \over \gamma_{l,l,2}} \\
			& & & & & \vec{\eta}^\top_{1} 
		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)}, \label{eqn:Dltdef}
}
where $\vec{\eta}_{0}, \vec{\eta}_{1} \in \R^{3(2n+1)}$ are vectors with entries given by 
\bseqn{
	\big(\vec{\eta}_{0}\big)_j &= 
		\begin{cases}
			{1 \over \alpha_{l,-l,3}} & j = 1 \\
			{- \; \alpha_{l,-l,4} \over \alpha_{l,-l,3} \: \gamma_{l,1-l,2}} & j = 2(2n+1) + 2 \\
			0 & o/w
		\end{cases}, \\
	\big(\vec{\eta}_{1}\big)_j &= 
		\begin{cases}
			{1 \over \alpha_{l,l,4}} & j = 2l+1 \\
			{- \; \alpha_{l,l,3} \over \alpha_{l,l,4} \: \gamma_{l,l-1,2}} & j = 3(2n+1) - 1 \\
			0 & o/w
		\end{cases}.
}
For $l = 0$ we can simply set
\bseqnnumber{
	D^T_0 = 
		\begin{pmatrix}
			{\beta_{0,0,4} \over \alpha_{0,0,3} \beta_{0,0,4} - \alpha_{0,0,4} \beta_{0,0,3}} & {- \alpha_{0,0,4} \over \alpha_{0,0,3} \beta_{0,0,4} - \alpha_{0,0,4} \beta_{0,0,3}} & 0 \\
			0 & 0 & {1 \over \gamma_{0,0,2}} \\
			{\beta_{0,0,3} \over \alpha_{0,0,4} \beta_{0,0,3} - \alpha_{0,0,3} \beta_{0,0,4}} & {- \alpha_{0,0,3} \over \alpha_{0,0,4} \beta_{0,0,3} - \alpha_{0,0,3} \beta_{0,0,4}} & 0
		\end{pmatrix}. \label{eqn:Dzerotdef}
}
It will be useful for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
\begin{definition}\label{def:clenshawmats}
	The matrices $-\Dlt (B_l-G_l(x,y,z))$, $\Dlt C_l$ for $l \in \No$ defined via \bsrefeqns{eqn:clenshawmats1}{eqn:Dzerotdef} are called the \textbf{recurrence coefficient matrices} for a given family of multidimensional orthogonal polynomials.
\end{definition}

@@

\bstodoinline{Is this para enough? coeffs/ transform}
In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations. Naturally, we of course need a way to obtain the coefficients $f_{l,m}$. The coefficients can be calculated via the integral
\bseqn{
	f_{l,m} = \int_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
}
using the orthonormality of the spherical harmonics. Methods to calculate these coefficients exist, known as spectral transforms, and are well established for the spherical harmonics (see e.g. \cite{mohlenkamp1999fast, suda2002fast}).

@@

Let's explain what we mean. Suppose $v : \Omega \to \C$ is a function, and we encounter the problem of finding the coefficients of the expansion for $v(x,y,z) f(x,y,z)$. We wish to therefore find an operator $V$ for $v$ so that
\bseqn{
	v(x,y,z) f(x,y,z) = v(x,y,z) \fvec^\top \SHvec(x,y,z) = (V \fvec)^\top \SHvec(x,y,z),
}
i.e. $V \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto v(x,y,z) f(x,y,z)$ in the spherical harmonic basis.

@@

%%% TANGENT SPACE ----- VECTOR SHs
\section{Vector spherical harmonics as orthogonal vectors in three variables}\label{Section:VSHs}

@@

As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
\bseqnnumber{
	\VSHvecfull = 
		\begin{pmatrix} 
			\VSHvecfull_0 \\ \VSHvecfull_1 \\ \vdots 
		\end{pmatrix}, \quad \text{where } \quad 
	\VSHvecfulll = 
		\begin{pmatrix} 
			(\gradY^{-l}_l)^\top \\ (\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top \\ (\gradpY^{l}_l)^\top
		\end{pmatrix} \in \C^{2(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHfulldef}
}

@@

			\mathcal{A}_{l,-l,3} & 0_2 & \mathcal{A}_{l,-l,4} & & \\

@@

			& & \mathcal{A}_{l,l,3} & 0_2 & \mathcal{A}_{l,l,4} \\

@@

			0_2 & \mathcal{A}_{l,-l,6} & \\
			\mathcal{A}_{l,-l+1,5} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{A}_{l,l-1,6} \\
			& & \mathcal{A}_{l,l,5} & 0_2

@@

			0 & A_{l,m,j}
		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad

@@

			- A_{l,m,j} & 0
		\end{pmatrix} \: \text{for } j = 5, 6.

@@

			\mathcal{B}_{l,-l,3} & 0_2 & \mathcal{B}_{l,-l,4} & & \\

@@

			& & \mathcal{B}_{l,l,3} & 0_2 & \mathcal{B}_{l,l,4} \\

@@

			0_2 & \mathcal{B}_{l,-l,6} & \\
			\mathcal{B}_{l,-l+1,5} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{B}_{l,l-1,6} \\
			& & \mathcal{B}_{l,l,5} & 0_2

@@

			0 & B_{l,m,j}
		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad

@@

			- B_{l,m,j} & 0
		\end{pmatrix} \: \text{for } j = 5, 6.

@@

			0 & 0 & \Gamma_{l,-l,2} \\
			& & & \Gamma_{l,-l,2} \\

@@

			& & & & & \Gamma_{l,l,2} \\
			& & & & & & \Gamma_{l,l,2}& 0 & 0 \\

@@

			0 & \Gamma_{l,-l,3} \\
			- \Gamma_{l,-l,3} & 0 \\

@@

			& & & 0 & \Gamma_{l,l,3} \\
			& & & - \Gamma_{l,l,3} & 0 \\

@@

			& \Gamma_{l,-l+1,1} \\

@@

			& & & & \Gamma_{l,l-1,1} \\

@@

Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
\bseqn{
	\uvec(x,y,z) = u_\varphi(x,y,z) \: \phivec + u_\theta(x,y,z) \: \thetavec, \quad (x,y,z) \in \Omega
}
for some scalar functions $u^\varphi, u^\theta$ where $\phivec, \thetavec$ are the standard non-radial unit basis vectors on the sphere defined by $\phivec := (\cos \theta \sin \varphi, \sin \theta \sin \varphi, \cos \varphi)^\top$, $\thetavec := (-\sin \theta, \cos \theta, 0)^\top$. Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}. We can then approximate this function by its expansion in the VSH basis, i.e. for some large enough $N \in \N$:
\bseqn{
	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big].
}
